Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-15 manuscript/manuscript.tex
incorporates some of Peter's suggestions

\author{}
\date{}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing

@@

Population models are important tools for predicting the impacts of
environmental change on species. But reconciling the scales at which
population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The major
hurdle is that most population models, at least for plant species, are
built using data from small, localized plots because parameterizing
traditional population models requires tracking the fates of
individuals. These models are difficult to scale up from the micro to
meso-scales because the fitted parameters do not fully represent the
spatial variation present at scales beyond that at which the data are
collected (Sæther et al. 2007). At the same time, most demographic data
is collected over short time spans. For example, the most common study
duration in the COMPADRE matrix population model database is 4 years and
only a few exceed 10 years (Salguero-Gómez et al. 2015). The constrained
spatio-temporal extent of most demographic datasets reflects the
difficulty of collecting such data, but those constraints limit our
ability to extrapolate population models. Thus, our ability to use
population models to predict the consequences of climate change is
limited when we rely on individual-level data.

That population models do not scale easily explains, in part, why
species distribution models (SDMs) have become a major tool for land
managers and conservation planners (see Guisan and Thuiller 2005 for a
review). SDMs typically rely on easy-to-collect presence/absence data
(but see Clark et al. 2014 for new methods) that allow researchers to
cover large spatial extents (e.g., Maiorano et al. 2013). Likewise,
predictors in SDMs tend to be remotely sensed products. Thus, it is
relatively straightforward to parameterize and project SDMs at landscape
and regional scales, the very scales at which land managers work.
However, the limitations of SDMs are well known (Elith and Leathwick
2009). It seems that what land managers really want are population
models at the landscape and regional scale, if only the ecological
community could provide them.

Aggregate measures of individual plant performance, such as those
typically collected as part of large-scale census efforts, offer an
alternative to detailed demographic data for modeling populations (Clark
and Bjørnstad 2004, Freckleton et al. 2011). Such population-level data
will never match the precision of individual-level data, but it is more
feasible to attain a broad coverage sample when collecting coarse-scale
data. This presents a difficult trade-off: on the one hand,
individual-level data leads to more reliable models; on the other hand,
population-level data leads to models that will produce less precise
predictions but can be applied over greater spatial and temporal
extents. An open question is how well models based on population-level
data compare to models based on individual-level data.

To date, relatively few studies have tried to model populations based on
data other than detailed individual-level data. An important exception
is an effort by Taylor and Hastings (2004) to model the population
growth rate of an invasive species to investigate the best strategies
for invasion control. They used a ``density-structured'' model where the
state variable is a discrete density state rather than a continuous
density measure. Building on this work, Freckleton et al. (2011) showed
that density-structured models compare well to continuous models in
theory, and Queenborough et al. (2011) showed the application of such
methods in a study on arable weeds. In particular, Queenborough et al.
(2011) provide empirical evidence that density-structured models are
capable of reproducing population dynamics, even if some precision is
lost when compared to fully continuous models. Thus, population models
based on coarse, population-level data show promise for producing
ecological forecasts at landscape and regional scales (Queenborough et
al. 2011). However, none of these models included environmental
covariates.

Basing population models on aggregated individual-level data in a
climate change context is hampered by the fact that it is individuals
that respond to climate, not populations (Clark et al. 2012). This fact
puts us in uneasy proximity to an ``ecological fallacy'' where one
deduces inference on the individual from statistical inference on the
group (Piantadosi et al. 1988). For example, individual plants may
respond positively to precipitation but a negative trend is observed at
the population level due to increased competition among plants as they
grow larger and consume more resources. Thus, it is important to ask the
question: Can aggregated data be used to detect climate signals of the
same sign and magnitude as individual-level data? If not, then building
population models with climate covariates on aggregated data will lead
to incorrect forecasts.

Here, we test the assumption that statistical and population models
based on aggregated data can detect climate signals as wells as models
based on individual-level data. We use a unique demographic dataset that
tracks the fates of individual plants from four species over 14 years to
build single-species population models, since those are often used tools
for ecological forecasts and climate vulnerability assessments. We first
fit population models with interannual variation in vital rates
explained, in part, by climate covariates. We then perturb the climate
covariates to test the sensitivities of species to climate change. By
doing these analyses using both individual and aggregated forms of the
same data we can directly compare the two types of models.

In general, we find that population models based on detailed demographic
data are more accurate and precise than models based on aggregated data.
Both types of models are able to detect climate signals, as evidenced by
the sensitivity of simulated equilbrium plant cover under a perturbed
climate scenario. But the two types of models produce inconsistent
forecasts, in some cases producing completely opposing predictions. This
leads us to conclude that, at least for these species at this location,
detailed demographic data is necessary to detect the ``right'' climate
signal. A worrying caveat to our work is that forecasts from both models
were very uncertain. It seems that even 14 years worth of demographic
data is not enough to produce meaningful forecasts when model
uncertainty is explicitly considered.

@@

Our demographic data comes from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The
dataset is freely available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011) , and interested readers should refer to the
metadata therein for a complete description. The site is about 800 m
above sea level and mean annual precipitation (1878-2009) is 334 mm,
with most annual precipitation falling from April through September. The
site is grass dominated and, for the purposes of our study, we focus on
the four most abundant graminoid species: \emph{Bouteloua gracilis}
(BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii}
(PASM), and \emph{Poa secunda} (POSE) (Fig. 1).

From 1932 to 1945 individual plants were identified and mapped annually
in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were
distributed in six pastures, each assigned a grazing treatment of light
(1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76
ha/aum) stocking rates (two pastures per treatment). In this analysis we
account for potential differences among the grazing treatments, but do
not focus on grazing$\times$climate interactions. The annual maps of the
quadrats were digitized and the fates of individual plants tracked and
extracted using a computer program. Daily climate data, which we
aggregated into climate variables of interest, are available for the
duration of the data collection period (1932 - 1945) from the Miles City
airport, Wiley Field, 9 km from the study site.

In this paper, we model populations based on two levels of data:
individual and quadrat (Fig. 2). The individual data is the ``raw''
data. For the quadrat level we data we simply sum individual areal cover
for each quadrat by species. This is equivalent to a perfect census of
quadrat percent cover, so we do not need to consider measurement error.
Based on these two datasets we can compare population models built using
individual level data and aggregated quadrat level data.

All R code and data necessary to reproduce our analysis is archived on
GitHub as release v1.0\footnote{\emph{Note to reviewers}: so that v1.0
  will be associated with the published version of the manuscript, we
  have released v0.1 to be associated with this review version.}
(\url{http://github.com/atredennick/MicroMesoForecast/releases}). That
stable release will remain static as a record of this analysis, but
subsequent versions may appear if we update this work.

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual level data we fit models for survival, growth, and
recruitment of new individuals for each species. At the quadrat level we
fit a single regression model for population growth. We describe the
statistical models separately since fitting the models required
different approaches. All models contain five climate covariate that we
chose \emph{a priori}: ``water year'' precipitation at \emph{t}-1
(lagppt); fall through spring precipitation at \emph{t}-1 and \emph{t}-2
(ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1
and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}
is the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$
size interactions. Climate $\times$ size interactions are for climate
main effects only, that is we do not include interactions among size and
interacting climate effects.

We fit all models using a hierarchical Bayesian approach. The models are
fully descibed in Appendix A, so here we focus on the main process and
the mode likelihood. For the likelihood models, $\textbf{y}^X$ is always
the relevant vector of observations for vital rate \emph{X}
($X = S, G, R, or P for survival, growth, recruitment, or population growth) For example, $\textbf{y}\^{}S\$
is a vector of 0s and 1s indicating whether a genet surives from
\emph{t} to \emph{t+1}, or not.

@@

The statistical approach used to model vital rates using aggregated data
depends on the type of data collected. In our case, and as is often the
case with census data, we have percent cover data (which can easily be
transformed to proportion data, of course). We first considered fitting
three vital rate models analagous to those we fit at the individual
level: one for probability of extirpation within a quadrat (analagous to
survival), one for cover change within a quadrat (analagous to growth),
and one for probability of colonization within a quadrat (analagous to
recruitment). However, within-quadrat extirpation and colonization
events were rare in our time series ($N=9$ and $N=10$, respectively
across all species). Given the broad spatial distribution of the
quadrats we are studying, it is safe to assume that these events are in
fact rare enough to be ignored for our purposes. So we constrained our
statistical modeling of vital rates at the population level to change in
percent cover within quadrats. For the remaining discussion of
statistical modeling we refer to proportion data, which is simply
percent data divided by 100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in a quadrat-based population model
the simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change
($G$) from time $t$ to $t+1$ is

@@

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. We recignize
such short MCMC chains may surprise those more familiar with other MCMC
samplers (i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally
efficient, which reduces the number of iterations needed to achieve
convergence. We assessed convergence visually and made sure scale
reduction factors for all parameters were less than 1.01. For the
purposes of including parameter uncertainty in our population models, we
saved the final 1,000 iterations from each of the three MCMC chains for
all parameters to be used as randomly drawn values during population
simulation. This step alleviates the need to reduce model parameters by
model selection since sampling from the full parameter space in the MCMC
ensures that if a parameter broadly overlaps zero, on average the effect
in the population models will also be near zero. We report the posterior
mean, standard deviation, and 95\% Bayesian Credible Intervals for every
parameter of each model for each species in Appendix B.

@@

With the posterior distribution of the vital rate statistical models in
hand, it is straightforward to simulate the population models. We used
an Integral Projection Model (IPM) to model populations based on
individual level data and an quadrat based version of an
individually-based model (Quadrat-Based Model, QBM) to model populations
based on quadrat level data. We describe each in turn.

@@

We use an environmentally stochastic IPM (Rees and Ellner 2009) that
includes the random year effects and the climate covariates from the
vital rate statistical models. But note that we can, and do for some
simulations, ignore the random year effects so that only the climate
effects can drive interannual variation. Our IPM follows the
specification of Chu and Adler (2015) where the population of species
\emph{j} is a density function $n(u_{j},t)$ giving the density of
sized-\emph{u} genets at time \emph{t}. Genet size is on the natural log
scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the
arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. So, the
density function for any size \emph{v} at time $t+1$ is

@@

which, said plainly, means we are calculating growth (\emph{G}) for
individuals that survive (\emph{S}) from time \emph{t} to \emph{t+1} and
adding in newly recruited (\emph{R}) individuals of an average sized
one-year-old genet for the focal species. Our stastical model for
recruitment (\emph{R}, described above) returns the number of new
recruit produced per quadrat. Following previous work (Adler et al.
2012, Chu and Adler 2015), we assume that fecundity increases linearly
with size
($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$)
to incorporate the recruitment function in the spatially-implicit IPM.

@@

Our quad-based model (QBM) perfectly mirrors its statistical description
(Eqs. 8-9). We use the same approach for drawing parameter values as
described for the IPM.

@@

To test each model's ability to forecast the population state we made
out of sample predictions using leave-one-year-out cross validation. For
both levels of modeling, we fit the vital rate models using observations
from all years except one, and then used those fitted parameters in the
population models to perform a one-step-ahead forecast for the year
whose observations were withheld from model fitting. Within each
observation year, several quadrats are sampled. So we made predictions
for each observed quadrat in the focal year initialized with cover the
previous year. Since we were making quadrat specific predictions we
incorporated the group effect on the intercept for both models. We
repeated this procedure for all 13 observation years, making 100
one-step-ahead forecasts for each quadrat-year combination with
parameter uncertainty included via randomdrawd from the MCMC chain as
described above. Random year effects were set to zero since year effects
cannot be assigned to unobserved years.

This model validation allowed us to compare accuracy and precision of
the two modeling approaches (individual-level versus population-level).
We first calculated the median predicted cover across the 100
simulations for each quadrat-year and then calculated the absolute error
as the difference between the observed cover for a given quadrat-year
and the median prediction. To arrive at mean absolute error (MAE), we
then averaged the absolute error within each species across the
quadrat-year specific errors. We use MAE as our measure of accuracy. To
measure precision we calculated the distance between the upper and lower
90th quantiles of the 100 predictions and averaged this value over
quadrat-years for each species.

@@

Our main goal in this paper is to see if models based on aggregate level
data are as sensitive to climate as models based on individual level
data. So, with our fitted and validated models in hand, we ran
simulations for each model type (IPM and QBM) under four climate
perturbation scenarios: (1) observed climate, (2) precipitation
increased by 1\%, (3) temperature increased by 1\%, and (4)
precipitation and temperature increased by 1\%. We ran the simulations
for 2,500 time steps, enough to estimate equilibrium cover after
discarding an initial 500 time steps as burn-in. Each simulation was run
under two parameter scenarios: (1) using mean parameter estimates and
(2) using randomly drawn parameters from the MCMC chain. We use (1) to
detect the overall sensitivity of equilibrium cover to climate, and we
use (2) to show the impact of model uncertainty on forecast precision.

As an effort to identify potential discrepencies between IPM and QBM
forecasts, we also ran simulations designed to quantify the
sensitivities of individual and combined vital rates to climate for the
IPM. Specifically, we ran simulations for the above climate scenarios,
but applied the perturbed climate covariates to survival, growth, and
recruitment vital rates individually and in pairwise combinations. This
allows us to isolate the vital rate(s) most sensitive to climate. For
this analysis, we used mean parameter estimates to reduce the sources of
uncertainty in the sensitivity estimates.

@@

@@

Our goal was not to make any explicit forecast for the future state of
these populations based on predicted climate change. But our results
highlight the state of affairs in ecology when it comes to forecasting
the impacts of climate change. The analysis we conducted here could be
considered, with some exceptions of course, at the forefront of
ecological forecasting in terms of the statistical approach employed
(hierarchical Bayesian), the type of population model we used
(stochastic IPM with parameter uncertainty), and the amount of high
quality data we had at our disposal (14 years of individual level data).
Yet, model predictions proved so uncertain that any forecast, when
bounded with model and parameter uncertainty, would be at best not
useful and at worst meaningless. How might we improve on this state of
affairs?

@@

@@

@@

>>>>

\author{}
\date{}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{todonotes}

@@

Population models are important tools for predicting the impacts of
environmental change on species persistence and abundance. But
reconciling the scales at which population models are parameterized and
the scales at which environmental changes play out remains a challenge
(Clark et al. 2010, 2012, Freckleton et al. 2011, Queenborough et al.
2011). The major hurdle is that most population models are built using
data from a single study site because collecting those data, which
involves tracking the fates of individuals plants, is so difficult. The
resulting models cannot be applied to the landscape and regional scales
relevant to decision-makers without information about how the fitted
parameters respond to spatial variation in biotic and abiotic drivers
(Sæther et al. 2007). The temporal coverage of demographic datasets is
no better. {[}I think the temporal extent is a separate issue, might
muddy the waters to bring it up here\ldots{}{]} For example, the most
common study duration in the COMPADRE matrix population model database
is 4 years and only a few studies exceed 10 years (Salguero-Gómez et al.
2015). The limited spatio-temporal extent of individual-level
demographic datasets constrains our ability to use population models to
address applied questions about the consequences of climate change.

The inability of many population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions, the scales at which many land-use decisions are
made. However, the limitations of SDMs are well known (Elith and
Leathwick 2009). Ideally, researchers would provide managers with
landscape-scale population models, combing the extent of SDMs with
information about dynamics and species abundances.

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bjørnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but it might
be sufficient for modeling future population states, especially since it
is feasible to collect across broad spatial extents (e.g., Queenborough
et al. 2011). The choice between individual and population-level data
involves a difficult trade-off: while individual-level data leads to
more mechanistic models, population-level data leads to models that can
be applied over greater spatial and temporal extents. An open question
is how much forecasting skill is lost when we build models based on
population rather than individual-level data.

To date, most empirical population modelers have relied on
individual-level data, with few attempts to exploit population-level
data. An important exception was an effort by Taylor and Hastings (2004)
to model the population growth rate of an invasive species to identify
the best strategies for invasion control. They used a
``density-structured'' model where the state variable is a discrete
density state rather than a continuous density measure. Building on this
work, Freckleton et al. (2011) showed that density-structured models
compare well to continuous models in theory, and Queenborough et al.
(2011) demonstrated the application of such methods in a study on arable
weeds. In particular, Queenborough et al. (2011) provide empirical
evidence that density-structured models are capable of reproducing
population dynamics, even if some precision is lost when compared to
fully continuous models. Thus, population models based on coarse,
population-level data show promise for producing ecological forecasts at
landscape and regional scales (Queenborough et al. 2011). However, none
of these models included environmental covariates.

Addressing climate change questions with models fit to population-level
data is potentially problematic. It is individuals, not populations,
that respond to climate variables (Clark et al. 2012). Ignoring this
fact puts us in uneasy proximity to an ``ecological fallacy'', where
inference about the individual relies on statistical inference on the
group (Piantadosi et al. 1988). Population growth (or decline) is an
epiphenomenon of demographic processes like survival, growth, and
recruitment that occur at the level of individual plants. Climate can
affect each demographic process in unique, potentially opossing, ways
(Dalgleish et al. 2011). These unique climate responses may difficult
resolve in statistical models based on population-level data where
demographic processes are not identifiable. If population-level data
cannot detect important impacts of climate drivers on populations, then
population models built with such data will make poor forecasts.

Here, we ask whether statistical and population models based on
aggregated, population-level data can detect climate signals as wells as
models based on individual-level data. We used a unique demographic
dataset that tracks the fates of individual plants from four species
over 14 years to build two kinds of single-species population models,
traditional models using individual growth, survival, and recruitment
data and alternative models based on basal cover. In both models,
interannual variation is explained, in part, by climate covariates. We
then performed simulations to quantify the sensitivities of species'
cover to small perturbations in average precipitation and temperature.
We found that population models based on detailed demographic data are
more accurate and precise than models based on aggregated data. Our
results suggest the popuation-level model is less accurate and precise
because important demographic climate signals go undetected. For these
species at this location, detailed demographic data appears necessary to
make accurate forecasts. A worrying caveat to our work is that forecasts
from both models were very uncertain when we considered full process and
parameter uncertainty. Even 14 years worth of demographic data may not
be sufficient to make useful forecasts.

@@

Our demographic data come from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The
dataset is freely available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011) , and interested readers should refer to the
metadata for a complete description. The site is about 800 m above sea
level and mean annual precipitation (1878-2009) is 334 mm, with most
annual precipitation falling from April through September. The community
is grass-dominated and we focused on the four most abundant grass
species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata}
(HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE)
(Fig. 1).

From 1932 to 1945 individual plants were identified and mapped annually
in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were
distributed in six pastures, each assigned a grazing treatment of light
(1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76
ha/aum) stocking rates (two pastures per treatment). In this analysis we
account for potential differences among the grazing treatments, but do
not focus on grazing$\times$climate interactions. The annual maps of the
quadrats were digitized and the fates of individual plants tracked and
extracted using a computer program (Lauenroth and Adler 2008, Chu et al.
2014). Daily climate data, which we aggregated into climate variables of
interest, are available for the duration of the data collection period
(1932 - 1945) from the Miles City airport, Wiley Field, 9 km from the
study site.

In this paper, we model populations based on two levels of data:
individual and quadrat (Fig. 2). The individual data is the ``raw''
data. For the quadrat level we data we simply sum individual basal cover
for each quadrat by species. This is equivalent to a near-perfect census
of quadrat percent cover because previous analysis shows that
measurement error at the individual level is small (Chu and Adler 2014).
Based on these two datasets we can compare population models built using
individual level data and aggregated quadrat level data.

All R code and data necessary to reproduce our analysis is archived on
GitHub as release v1.0\footnote{\emph{Note to reviewers}: so that v1.0
  will be associated with the published version of the manuscript, we
  have released v0.1 to be associated with this review version.}
(\url{http://github.com/atredennick/MicroMesoForecast/releases}). That
stable release will remain static as a record of this analysis, but
subsequent versions may appear if we update this work. We have also
deposited the v1.0 release on Dryad (\emph{link here after acceptance}).

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual-level data we fit models for survival, growth, and
recruitment for each species. At the quadrat-level we fit a single
regression model for population growth. We describe the statistical
models separately since fitting the models required different
approaches. All models contain five climate covariate that we chose
\emph{a priori}: ``water year'' precipitation at \emph{t}-1 (lagppt);
fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and mean spring temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$
size interactions. Climate $\times$ size interactions are for climate
main effects only, that is we do not include interactions among size and
interacting climate effects.

\todo{let's talk about timing of "spring". Miles City gets a lot of moisture in May and June...}

We fit all models using a hierarchical Bayesian approach. The models are
fully descibed in Appendix A, so here we focus on the main process and
the model likelihood. For the likelihood models, $\textbf{y}^X$ is
always the relevant vector of observations for vital rate \emph{X}
($X = S, G, R, or P for survival, growth, recruitment, or population growth). For example, $\textbf{y}\^{}S\$
is a vector of 0s and 1s indicating whether a genet survives from
\emph{t} to \emph{t+1}, or not.

@@

The statistical approach used to model vital rates using aggregated data
depends on the type of data collected. In our case, and as is often the
case with census data, we have percent cover data (which can easily be
transformed to proportion data). We first considered fitting three vital
rate models analagous to those we fit at the individual level: one for
probability of extirpation within a quadrat (analagous to survival), one
for cover change within a quadrat (analagous to growth), and one for
probability of colonization within a quadrat (analagous to recruitment).
However, within-quadrat extirpation and colonization events were rare in
our time series ($N=9$ and $N=10$, respectively, across all species).
Given the broad spatial distribution of the quadrats we are studying, it
is safe to assume that these events are in fact rare enough to be
ignored for our purposes. So we constrained our statistical modeling of
vital rates at the population level to change in percent cover within
quadrats. For the remaining discussion of statistical modeling we refer
to proportion data, which is simply percent data divided by 100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in a quadrat-based population model,
the simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change
($G$) from time $t$ to $t+1$ is

@@

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. We recignize
such short MCMC chains may surprise those more familiar with other MCMC
samplers (i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally
efficient, which reduces the number of iterations needed to achieve
convergence. We assessed convergence visually and made sure scale
reduction factors for all parameters were less than 1.01. For the
purposes of including parameter uncertainty in our population models, we
saved the final 1,000 iterations from each of the three MCMC chains to
be used as randomly drawn values during population simulation. This step
alleviates the need to reduce model parameters by model selection since
sampling from the full parameter space in the MCMC ensures that if a
parameter broadly overlaps zero, on average the effect in the population
models will also be near zero. We report the posterior mean, standard
deviation, and 95\% Bayesian Credible Intervals for every parameter of
each model for each species in Appendix B.

@@

With the posterior distribution of the vital rate statistical models in
hand, it is straightforward to simulate the population models. We used
an Integral Projection Model (IPM) to model populations based on
individual level data {[}cite Ellner and Rees 2006{]} and an quadrat
based version of an individually-based model (Quadrat-Based Model, QBM)
to model populations based on quadrat level data. We describe each in
turn.

@@

We use an environmentally stochastic IPM (Rees and Ellner 2009) that
includes the random year effects and the climate covariates from the
vital rate statistical models. However, for some simulations, we ignore
the random year effects so that only the climate effects drive
interannual variation. Our IPM follows the specification of Chu and
Adler (2015) where the population of species \emph{j} is a density
function $n(u_{j},t)$ giving the density of sized-\emph{u} genets at
time \emph{t}. Genet size is on the natural log scale, so that
$n(u_{j},t)du$ is the number of genets whose area (on the arithmetic
scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. So, the density
function for any size \emph{v} at time $t+1$ is

@@

which means we are calculating growth (\emph{G}) for individuals that
survive (\emph{S}) from time \emph{t} to \emph{t+1} and adding in newly
recruited (\emph{R}) individuals of an average sized one-year-old genet
for the focal species. Our stastical model for recruitment (\emph{R},
described above) returns the number of new recruit produced per quadrat.
Following previous work (Adler et al. 2012, Chu and Adler 2015), we
assume that fecundity increases linearly with size
($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$)
to incorporate the recruitment function in the spatially-implicit IPM.
{[}You probably should say something about spatial structure and cite
Adler et al. 2010 EcoLetts for the ``no overlap rule''{]}

@@

Our quad-based model (QBM) perfectly mirrors {[}I think you need to say
a little more about how you simulate it{]} its statistical description
(Eqs. 8-9). We use the same approach for drawing parameter values as
described for the IPM.

@@

To test each model's ability to forecast population state, we made
out-of-sample predictions using leave-one-year-out cross validation. For
both levels of modeling, we fit the vital rate models using observations
from all years except one, and then used those fitted parameters in the
population models to perform a one-step-ahead forecast for the year
whose observations were withheld from model fitting. Within each
observation year, several quadrats were sampled. So we made predictions
for each observed quadrat in the focal year, initializing each
simulation with cover in the quadrat the previous year. Since we were
making quadrat-specific predictions, we incorporated the group effect on
the intercept for both models. We repeated this procedure for all 13
observation years, making 100 one-step-ahead forecasts for each
quadrat-year combination with parameter uncertainty included via random
draw from the MCMC chain as described above. Random year effects were
set to zero since year effects cannot be assigned to unobserved years.

This model validation allowed us to compare accuracy and precision of
the two modeling approaches (IPM versus QBM). We first calculated the
median predicted cover across the 100 simulations for each quadrat-year
and then calculated the absolute error as the absolute value of the
difference between the observed cover for a given quadrat-year and the
median prediction. To arrive at mean absolute error (MAE), we then
averaged the absolute error within each species across the quadrat-year
specific errors. We use MAE as our measure of accuracy. To measure
precision we calculated the distance between the upper and lower 90th
quantiles of the 100 predictions and averaged this value over
quadrat-years for each species.

@@

Our main goal in this paper is to see if models based on aggregated data
are as sensitive to climate covariates as models based on individual
level data. So, with our fitted and validated models in hand, we ran
simulations for each model type (IPM and QBM) under four climate
perturbation scenarios: (1) observed climate, (2) precipitation
increased by 1\%, (3) temperature increased by 1\%, and (4)
precipitation and temperature increased by 1\%. We ran the simulations
for 2,500 time steps, enough to estimate equilibrium cover after
discarding an initial 500 time steps as burn-in. Each simulation was run
under two parameter scenarios: (1) using mean parameter estimates and
(2) using randomly drawn parameters from the MCMC chain. We use (1) to
detect the overall sensitivity of equilibrium cover to climate, and we
use (2) to show the impact of model uncertainty on forecast precision.

As an effort to identify potential discrepencies between IPM and QBM
forecasts, we also ran simulations designed to quantify the
sensitivities of individual and combined vital rates to climate for the
IPM. Specifically, we ran simulations for the above climate scenarios,
but applied the perturbed climate covariates to survival, growth, or
recruitment vital rates individually and in pairwise combinations. This
allowed us to isolate the vital rate(s) most sensitive to climate. For
this analysis, we used mean parameter estimates to reduce the sources of
uncertainty in the sensitivity estimates.

@@

MOVE FIRST TWO PARAGRAPHS TO INTRO. MOVE FIG 5 and 6 TO METHODS AND
RESULTS

@@

Our goal was not to make any explicit forecast for the future state of
these populations based on predicted climate change. But our results
highlight the state of affairs in ecology when it comes to forecasting
the impacts of climate change. The analysis we conducted here could be
considered, with some exceptions of course, at the forefront of
ecological forecasting in terms of the statistical approach employed
(hierarchical Bayesian), the type of population model we used
(stochastic IPM with parameter uncertainty), and the amount of high
quality data we had at our disposal (14 years of individual level data).
Yet, model predictions proved so uncertain that any forecast, when
bounded with model and parameter uncertainty, would be at best not
useful and at worst meaningless. For all species, the 90\% quantiles of
predicted changes in populatoin size overlapped zero; we cannot even
predict whether a change in precipitation or temperature will cause
populations to increase or decrease. How might we improve on this state
of affairs?

@@

Chu, C., K. M. Havstad, N. Kaplan, W. K. Lauenroth, M. P. McClaran, D.
P. Peters, L. T. Vermeire, and P. B. Adler. 2014. Life form influences
survivorship patterns for 109 herbaceous perennials from six semi-arid
ecosystems. Journal of Vegetation Science 25:947--954.

@@

Dalgleish, H. J., D. N. Koons, M. B. Hooten, C. A. Moffet, and P. B.
Adler. 2011. Climate influences the demography of three dominant
sagebrush steppe plants. Ecology 92:75--85.

@@

Lauenroth, W. K., and P. B. Adler. 2008. Demography of perennial
grassland plants: Survival, life expectancy and life span. Journal of
Ecology 96:1023--1032.
